---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Anton Szimmat (12797237)"
date: "`r Sys.Date()`"
output: distill::distill_article
editor_options: 
  markdown: 
    wrap: 72
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers
    (Matrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your
    results.
-   Use a level 2 heading for each new exercise and answer each subtask
    next to its bullet points or use a new level 3 heading if you want.
-   Always render the R code for your solutions (`echo=TRUE`) and make
    sure to include the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless
        specifically instructed to).
-   Always submit both the rendered document(s) as well as your source
    Rmarkdown or Quarto document. Submit the files separately on moodle,
    **not** as a zip archive.
-   Submission format is HTML. Other formats will lead to a deduction of
    points.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `2025-exeRcise-sheet-3`. If you forgot how to do this, you can
    follow this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)
    -   Two Strengths of git

        -   Git is really great for collaboration with other users and
            if you want to understand the changes done by other users

        -   Git prevents the user from deleting code or other changes on
            a files by accident

    -   Two Weaknesses of git

        -   Git often struggles with binary (i.e. non-text) files

        -   Git usually struggles with large files (e. g. huge datasets)
e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves. You
    can, but don't need to create a `.gitignore` file. Just do not
    commit these files manually.
    -   The step in which the source md-document was knitted to html was
        saved in Git with the commit message "Knitting the Template with
        exercise 1 finished and using the distill package to html"

## Exercise 2: Putting your Repository on GitHub (3 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:
    -   The link to the repository is provided
        [here](https://github.com/ATSzimmat/exeRcise-sheet-3)
b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.
    -   Test change that can be pushed to GitHub

## Exercise 3: Pixar Films (4 Points)

Download the `pixar_films` and `public_response` datasets from the
GitHub repository and track them in git.

Link:
<https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-03-11>

For small datasets like these adding them to git is not a problem.

a)  Load the `pixar_films` dataset into R. Clean the dataset by removing
    films without a title. Inspect the variable `film_rating`. What are
    the possible values and what do they mean? Create a factor variable
    for the film rating. Why is this appropriate?

```{r}
# Loading the dataset into R
library(tidyverse)
pixar_data <- read_csv("data/pixar_films.csv")

# Revoming every film without a title
pixar_data <- pixar_data %>% filter(!is.na(film))

# Inspecting the variable film_rating
pixar_data$film_rating
```

-   The variable film_rating shows the suitability of a film for certain
    audiences based on its content. The possible values are based on the
    Motion Picture Association (MPA) film rating system.

-   There are 5 possible ratings in which the films can be classified
    (source: <https://www.filmratings.com/>)

    -   "G" (General Audiences) means that all are ages admitted

    -   "PG" (Parental Guidance Suggested) means that there is some
        material in the film that may not be suitable for children

    -   "PG-13" (Parents Strongly Cautioned) means that some material in
        the film may be inappropriate for children under 13

    -   "R" (Restricted) means that *under 17 requires accompanying
        parent or adult guardian.* The film contains some adult
        material. Parents are urged to learn more about the film before
        taking their young children with them

    -   "NC-17" (Adults Only) means that the film is suitable only for
        persons older than 17 years . Children are not admitted

```{r}
# Checking the type of the variable
typeof(pixar_data$film_rating)
# Creating a factor variable for the variable
pixar_data$film_rating <- as.factor(pixar_data$film_rating)
# Checking the type of the variable again
typeof(pixar_data$film_rating)


```

-   This is appropriate because in R a factor variable has a fixed set
    of possible values and the film rating variable is based on the MPA
    film rating system which was such afixed set of possible values

Inspect the film titles manually. Which films form a film series? A film
series can be identified by a common word in the titles of the films,
often in conjunction with a number in the title, e.g. "Despicable Me"
and "Despicable Me 2". Create a dataframe which displays a list of the
different series with the titles of the films and how many films belong
to the series. Output the dataframe.

```{r}
# Loading the tidyverse
library(tidyverse)
# Inspecting the variable film titles
pixar_data$film
```

The following films form a series

-   Cars (Cars, Cars 2, Cars 3)

-   Finding Nemo (Finding Nemo, Finding Dory (because Finding Dory is a
    Sequel to Finding Nemo))

-   Insight out

-   Lightyear

-   Monsters, Inc. (Monsters, Inc., Monsters University (because
    Monsters University is a Prequel to Monsters, Inc.))

-   The Incredibles (The Incredibles, Incredibles 2)

-   Toy Story (Toy Story, Toy Story 2, Toy Story 3, Toy Story 4,
    Lightyear (because Lightyear is a spin-off of the Toy Story series))

```{r}
# Creating the requested dataframe

# Creating a new variable that displays all the film titles without numbers; also I removed the article "the" from the film names because otherwhise it'll cause troubles
pixar_data_II <- pixar_data%>%
  mutate(film_series = str_remove(film, "^(The)\\s"),
film_series = str_remove(film_series, "\\s\\d+$"))

# Creating a dataframe that displays a list of the different series with the titles of the films and how many films belong to the series
pixar_data_II <- pixar_data_II %>%
  group_by(film_series) %>%
  summarise(
    films = paste(film, collapse = ", "),
    count = n())


# Now the article "the" is added again to two film series
pixar_data_II <- pixar_data_II %>%
                 mutate(film_series = case_when(film_series == "Incredibles" ~
                                  "The Incredibles",
                                  film_series == "Good Dinosaur" ~ 
                                  "The Good Dinosaur", TRUE ~ film_series)) 

# Now a small correction ...
pixar_data_II <- pixar_data_II %>% 
                 mutate(films = case_when(films == "Finding Nemo" ~
  "Finding Nemo, Finding Dory",
  films == "Monsters, Inc." ~ 
  "Monsters, Inc., Monsters University", 
  films == "Toy Story, Toy Story 2, Toy Story 3, Toy Story 4" ~
  "Toy Story, Toy Story 2, Toy Story 3, Toy Story 4, Lightyear",
  TRUE ~ film_series)) %>% slice(-c(5, 10, 12))

# Output the new dataframe
pixar_data_II
```

-   Load the `public_response` dataframe into R. Convert the
    `cinema_score` variable into a factor while ensuring the factor
    levels are defined in ascending order, from the lowest to the
    highest score. Combine `public_response` with the `pixar_films`
    dataset using an appropriate merge variable.

-   Choose one of the variables representing the public response and
    create a bar plot for the films belonging to a series. Here are the
    details of the plot:

    -   The film series are represented on the x-axis.
    -   Your chosen public response variable is displayed on the y-axis.
    -   Each film in the series is represented as a separate bar. Bars
        are grouped by film under their respective series on the x-axis.
        Order the bars within a series according to the release date of
        the films.
    -   A title and axis labels for context.

    What do you notice when comparing the scores of the films in a
    series? Do you see any patterns?

## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from
[Our World in Data](https://ourworldindata.org/) and analyze it, while
determining the research question yourself.

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets>
    and choose a dataset that interests you. You can have a look at
    <https://ourworldindata.org/> to gather some inspiration.
b)  Download the dataset and track it in git.
c)  Put the name / title of the dataset and a link to it below.

-   Dataset Name: ...
-   Link: <https://github.com/owid/owid-datasets/>...

d)  Come up with a (research) question you want to answer with the data
    and briefly explain why you believe this is an interesting question
    within one sentence. It should be a question that can be answered
    with the dataset and using R.
e)  Use R to answer your chosen question. Please limit your analysis to
    the functions and techniques we have covered so far in the course.
    You are **not expected** to use advanced statistical models or
    external packages which haven't been introduced.
f)  Create a meaningful plot / figure with the dataset. Make sure to
    provide a figure caption (via the chunk options / Rmarkdown) and
    correctly label the figure.

## Final Note

Make sure to push all your commits and changes to GitHub before
submitting the exercise sheet.
